Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States

### 2017-10-31 doc/manuscript/manuscript.Rmd
address L58--67 to a degree, but probably need more finess
alphabetize state names to address #15

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

Nearly all population genetic studies of *S. sclerotiorum* employ a macroscopic assay to determine mycelial compatibility, the ability for fungal hyphae from different colonies to appear to grow together without forming a barrier of dead cells between them (known as a barrage line) [@leslie1993fungal; @sirjusingh2001characterisation].
Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].

In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S. sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia. 
We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
As these nurseries were not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity of the pathogen by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus haplotype and mycelial compatibility group.
Knowing what variables predict genetic structure can help direct breeding efforts.
By investigating these aims, we hope to effectively describe the population structure of *S. sclerotiorum* in the USA and make available our database of isolates for use in future dry bean breeding efforts.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)). 
After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCl, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days. 
An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
We observed a total of 165 MLHs (215 with 16 loci).
Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig \@ref(biggraph)A,C).

@@

We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph), Table \@ref(tab:mlg-table)). 
Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Although analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
Relative to the US, the international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.

@@

The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
While seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management], it has since been shown that *S. sclerotiorum* infections can be transmitted through seed [@botelho2013performance].
Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between France, Australia, and the US.
While we speculate that these transmission events are rare due to the genetic structuring by Region, these results suggest that seedborne infections may indeed reflect a source of inoculum.
This may, in turn increase the risk of introducing new sources of genetic variation through potential outcrossing events.

When we tested for sexual reproduction, we were unable to find evidence for it in any region except for CA and Australia. 
While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].
The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but we can see in Figure \@ref(DAPC-RY) that there is differentiation by year. 
Thus, this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].

>>>>

Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use]. 
These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
The genetic structure and mode of reproduction in these populations, however, is currently unknown.

@@

Nearly all population genetic studies of *S. sclerotiorum* employ a macroscopic assay to determine mycelial compatibility, the ability for fungal hyphae from different colonies to appear to grow together without forming a barrier of dead cells between them (known as a barrage line) [@leslie1993fungal; @sirjusingh2001characterisation].
Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
However, a question remains about the relationship between mycelial compatibility groups and clonal genotypes.

In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S. sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the Australia, France, Mexico, and the United States. 
We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
As these nurseries were not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity of the pathogen by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus haplotype and mycelial compatibility group.
Knowing what variables predict genetic structure can help direct breeding efforts.
By investigating these aims, we hope to effectively describe the population structure of *S. sclerotiorum* in the USA and make available our database of isolates for use in future dry bean breeding efforts.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the Australia, France, Mexico, and the United States (Table \@ref(tab:isolate-table)). 
After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCl, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days. 
An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the United States, as well as Australia, France, and Mexico (Table \@ref(tab:isolate-table)).
We observed a total of 165 MLHs (215 with 16 loci).
Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig \@ref(biggraph)A,C).

@@

We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph), Table \@ref(tab:mlg-table)). 
Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, ND, NE, MI), and an international community (Australia, France, MN).
Although analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
Relative to the US, the international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.

@@

The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
While seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management], it has since been shown that *S. sclerotiorum* infections can be transmitted through seed [@botelho2013performance].
Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between Australia, France, and the United States.
While we speculate that these transmission events are rare due to the genetic structuring by Region, these results suggest that seedborne infections may indeed reflect a source of inoculum.
This may, in turn increase the risk of introducing new sources of genetic variation through potential outcrossing events.

When we tested for sexual reproduction, we were unable to find evidence for it in any region except for Australia and CA. 
While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].
The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but we can see in Figure \@ref(DAPC-RY) that there is differentiation by year. 
Thus, this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
